Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-09-26 arxiv/proof_same_cov.tex
wip

	\paragraph{Proof of Claim \eqref{lem_exp_opt}.}
	To facilitate the analysis, we consider the following matrix notations.
	Let
		\[ \cE A^{\top} = {\sum_{k=1}^t \varepsilon^{(k)} A_k^{\top}}, \text{ and } \cW \define X(X^{\top} X)^{-1} X^{\top} \cE A^{\top} (AA^{\top})^{-1}. \]
	For any $j = 1,2,\dots, t$, let
	\begin{align*}
		H_j &\define B^{\star} A^{\top} (AA^{\top})^{-1} A_j - \beta^{(j)}, \\
		E_j &\define \cW A_j - \varepsilon^{(j)}.
	\end{align*}
	Then, we can write the function $g(A)$ conveniently as
	\[ g(A) = \sum_{j=1}^t \bignorm{X H_j + E_j}^2. \]
	We will divide $g(A)$ into three parts.
	First, we have that
	\begin{align}\label{eq_gA_p1}
		\sum_{j=1}^t \bignorm{X H_j}^2 = \bignorm{X (B^{\star} U_A U_A^{\top} - B^{\star})}^2,
	\end{align}
	whose expectation over $X$ is $\bignorm{\Sigma^{1/2}(B^{\star} U_A U_A^{\top} - B^{\star})}^2$.
	Second, one can verify that the cross terms are equal to the following
	\begin{align}\label{eq_gA_p2}
		\sum_{j=1}^t\inner{XH_j}{E_j} = \inner{X(B^{\star} U_A U_A^{\top} - B^{\star})}{\cW A - \cE}
		= - \inner{X (B^{\star} U_A U_A^{\top} - B^{\star})}{\cE},
	\end{align}
	which is zero in expectation over $\cE$.
	Finally, we have that
	\begin{align}\label{eq_gA_p3}
		\sum_{j=1}^t \norm{E_j}^2 &= \bignorm{\cW A - \cE}^2
		= \bignorm{\cE}^2 - \inner{\cW A}{\cE},
	\end{align}
	which is because $\norm{\cW A}^2 = \inner{\cW A}{\cE}$.
	Hence, it suffices to show that the expectation of equation \eqref{eq_gA_p3} is equal to $\sigma^2 (n\cdot t - p\cdot r)$.
	First, we have that $\ex{\cE} = \sigma^2 \cdot n \cdot t$.
%	Conditional on $X$, we have that
%	\begin{align*}
%		\exarg{\cE}{\norm{E_j}^2}
%		= \exarg{\cE}{\sum_{j=1}^t \bigbrace{\bignorm{\cW A_j}^2 - 2\inner{\cW A_j}{\varepsilon^{(j)}} }} + \sigma^2 \cdot n \cdot t.
%	\end{align*}
%	We consider both terms in the above equation one by one.
	Second, note that
	%\begin{align}
	%	\sum_{j=1}^t \norm{\cW A_j}^2 = \sum_{j=1}^t \bigtr{\cW A_j A_j^{\top} \cW^{\top}} = \bigtr{\cW AA^{\top} \cW^{\top}} = \bigtr{U_X U_X^{\top} \cE A^{\top} (AA^{\top})^{-1} A \cE^{\top}}. \label{eq_proof_same_cov_1}
	%\end{align}
	%We observe that
	\begin{align*}
			\exarg{\cE}{\cE A^{\top} (AA^{\top})^{-1} A \cE^{\top}}
		&= \exarg{\cE}{\sum_{j=1}^t \sum_{k=1}^t \varepsilon^{(j)} A_j^{\top} (AA^{\top})^{-1} A_k {\varepsilon^{(k)}}^{\top}} \\
		&= \exarg{\cE}{\sum_{j=1}^t \varepsilon^{(j)} A_j^{\top} (AA^{\top})^{-1} A_j {\varepsilon^{(j)}}^{\top}} \\
		&= \sigma^2 \cdot r \cdot \id_{n\times n}.
	\end{align*}
	In the above, the first equation uses the fact that for any $j\neq k$, $\varepsilon^{(j)}$ and $\varepsilon^{(k)}$ are pairwise independent.
	The second equation uses the fact that $\sum_{j=1}^t A_j^{\top} (AA^{\top})^{-1} A_j = \bigtr{\id_{r\times r}} = r$, and $\ex{\varepsilon^{(j)} {\varepsilon^{(j)}}^{\top}} = \sigma^2 \cdot \id_{n\times n}$.
	Therefore, we have that
	\begin{align*}
		\inner{\cW A}{\cE} = \sigma^2 \cdot r \cdot \bigtr{U_X U_X^{\top}} = \sigma^2 \cdot r \cdot p,
	\end{align*}
	because $U_X$ has rank $p$ by Fact \ref{lem_minv}.
	%Next, note that
	%\[ \sum_{j=1}^t \inner{\cW A_j}{\varepsilon^{(j)}} = \bigtr{\cW A\cE^{\top}}, \]
	%which is equal to equation \eqref{eq_proof_same_cov_1} by the definition of $\cW$.
	Hence the proof is complete.

	\paragraph{Proof of Claim \ref{claim_opt_dist}.}
	In order to show that the subspace spanned by the rows of $\hat{A}$ is close to $U_r$, we first show that $g(A)$ and its expectation is close as follows.
	\begin{align}\label{eq_gA_err}
		\bigabs{g(A) - \exarg{\cE, X}{g(A)}} \lesssim p^{-c} \cdot n \bignorm{\Sigma^{1/2} B^{\star} (U_AU_A^{\top} - \id)}^2 + p^{-1/2 + e} \cdot \sigma^2 \cdot n \cdot t.
	\end{align}
	To show the above result, we consider the concentration error of each part of $g(A)$.

	For equation \eqref{eq_gA_p1}, % $g_0(\cal W)= \sum_{j=1}^t\left\|Z v_j \right\|^2$.
%	$$g_0(\cal W)= \sum_{j=1}^t\left\|Z v_j \right\|^2,\quad v_j:= \Sigma^{1/2}\left(B^{\star} \cW^{\top} (\cW\cW^{\top})^{-1} W_j -  \beta_j\right).$$
	note that $X H_j = Z \Sigma^{1/2} H_j \in \real^n$ is a random vector with i.i.d. entries of mean zero, variance $\|\Sigma^{1/2}H_j\|^2$, and finite $\varphi$-th moment by Assumption \eqref{assume_rm}.
	Hence by the law of large numbers, we have that with high probability
	${\left\|Z v_j \right\|^2 } = { n \|v_j\|^2} \cdot (1 + \OO(p^{-c}))$,
	which implies that
	\begin{align}
		\abs{\sum_{j=1}^t \|X H_j\|^2 - \exarg{X}{\sum_{j=1}^t \norm{X H_j}^2}} \lesssim p^{-c} \cdot n \bignorm{\Sigma^{1/2} B^{\star} (U_{A}U_A^{\top} - \id_{t\times t})}^2. \label{eq_gA_err1}
	\end{align}

@@

		\abs{\norm{\cE}^2 - \ex{\norm{\cE}^2}} \le p^{-1/2 + \e} \ex{\norm{\cE}^2}. \label{eq_gA_err3}

@@

		&\le p^{-c} \cdot n \bignormFro{\Sigma^{1/2} B^{\star}}^2 + p^{-1/2 + e}\cdot \sigma^2 n \cdot t. \label{eq_g_gap}

>>>>

	\paragraph{Proof of Claim \eqref{lem_exp_opt}.}
	To facilitate the analysis, we consider the following matrix notations.
	Let
		\[ \cE A^{\top} = {\sum_{k=1}^t \varepsilon^{(k)} A_k^{\top}}, \text{ and } \cW \define X(X^{\top} X)^{-1} X^{\top} \cE A^{\top} (AA^{\top})^{-1}. \]
	For any $j = 1,2,\dots, t$, let
	\begin{align*}
		H_j &\define B^{\star} A^{\top} (AA^{\top})^{-1} A_j - \beta^{(j)}, \\
		E_j &\define \cW A_j - \varepsilon^{(j)}.
	\end{align*}
	Then, we can write the function $g(A)$ conveniently as
	\[ g(A) = \sum_{j=1}^t \bignorm{X H_j + E_j}^2. \]
	We will divide $g(A)$ into three parts.
	First, we have that
	\begin{align}\label{eq_gA_p1}
		\sum_{j=1}^t \bignorm{X H_j}^2 = \bignorm{X (B^{\star} U_A U_A^{\top} - B^{\star})}^2,
	\end{align}
	whose expectation over $X$ is $\bignorm{\Sigma^{1/2}(B^{\star} U_A U_A^{\top} - B^{\star})}^2$.
	Second, one can verify that the cross terms are equal to the following
	\begin{align}\label{eq_gA_p2}
		\sum_{j=1}^t\inner{XH_j}{E_j} = \inner{X(B^{\star} U_A U_A^{\top} - B^{\star})}{\cW A - \cE}
		= - \inner{X (B^{\star} U_A U_A^{\top} - B^{\star})}{\cE},
	\end{align}
	which is zero in expectation over $\cE$.
	Finally, we have that
	\begin{align}\label{eq_gA_p3}
		\sum_{j=1}^t \norm{E_j}^2 &= \bignormFro{\cW A - \cE}^2
		= \bignormFro{\cE}^2 - \inner{\cW A}{\cE},
	\end{align}
	which is because $\norm{\cW A}^2 = \inner{\cW A}{\cE}$.
	Hence, it suffices to show that the expectation of equation \eqref{eq_gA_p3} is equal to $\sigma^2 (n\cdot t - p\cdot r)$.
	First, we have that $\ex{\cE} = \sigma^2 \cdot n \cdot t$.
%	Conditional on $X$, we have that
%	\begin{align*}
%		\exarg{\cE}{\norm{E_j}^2}
%		= \exarg{\cE}{\sum_{j=1}^t \bigbrace{\bignorm{\cW A_j}^2 - 2\inner{\cW A_j}{\varepsilon^{(j)}} }} + \sigma^2 \cdot n \cdot t.
%	\end{align*}
%	We consider both terms in the above equation one by one.
	Second, note that
	%\begin{align}
	%	\sum_{j=1}^t \norm{\cW A_j}^2 = \sum_{j=1}^t \bigtr{\cW A_j A_j^{\top} \cW^{\top}} = \bigtr{\cW AA^{\top} \cW^{\top}} = \bigtr{U_X U_X^{\top} \cE A^{\top} (AA^{\top})^{-1} A \cE^{\top}}. \label{eq_proof_same_cov_1}
	%\end{align}
	%We observe that
	\begin{align*}
			\exarg{\cE}{\cE A^{\top} (AA^{\top})^{-1} A \cE^{\top}}
		&= \exarg{\cE}{\sum_{j=1}^t \sum_{k=1}^t \varepsilon^{(j)} A_j^{\top} (AA^{\top})^{-1} A_k {\varepsilon^{(k)}}^{\top}} \\
		&= \exarg{\cE}{\sum_{j=1}^t \varepsilon^{(j)} A_j^{\top} (AA^{\top})^{-1} A_j {\varepsilon^{(j)}}^{\top}} \\
		&= \sigma^2 \cdot r \cdot \id_{n\times n}.
	\end{align*}
	In the above, the first equation uses the fact that for any $j\neq k$, $\varepsilon^{(j)}$ and $\varepsilon^{(k)}$ are pairwise independent.
	The second equation uses the fact that $\sum_{j=1}^t A_j^{\top} (AA^{\top})^{-1} A_j = \bigtr{\id_{r\times r}} = r$, and $\ex{\varepsilon^{(j)} {\varepsilon^{(j)}}^{\top}} = \sigma^2 \cdot \id_{n\times n}$.
	Therefore, we have that
	\begin{align*}
		\inner{\cW A}{\cE} = \sigma^2 \cdot r \cdot \bigtr{U_X U_X^{\top}} = \sigma^2 \cdot r \cdot p,
	\end{align*}
	because $U_X$ has rank $p$ by Fact \ref{lem_minv}.
	%Next, note that
	%\[ \sum_{j=1}^t \inner{\cW A_j}{\varepsilon^{(j)}} = \bigtr{\cW A\cE^{\top}}, \]
	%which is equal to equation \eqref{eq_proof_same_cov_1} by the definition of $\cW$.
	Hence the proof is complete.

	\paragraph{Proof of Claim \ref{claim_opt_dist}.}
	In order to show that the subspace spanned by the rows of $\hat{A}$ is close to $U_r$, we first show that $g(A)$ and its expectation is close as follows.
	\begin{align}\label{eq_gA_err}
		\bigabs{g(A) - \exarg{\cE, X}{g(A)}} \lesssim p^{-c_{\varphi}} \cdot n \bignorm{\Sigma^{1/2} B^{\star} (U_AU_A^{\top} - \id)}^2 + p^{-c_{\infty}} \cdot \sigma^2 \cdot n \cdot t.
	\end{align}
	To show the above result, we consider the concentration error of each part of $g(A)$.

	For equation \eqref{eq_gA_p1}, % $g_0(\cal W)= \sum_{j=1}^t\left\|Z v_j \right\|^2$.
%	$$g_0(\cal W)= \sum_{j=1}^t\left\|Z v_j \right\|^2,\quad v_j:= \Sigma^{1/2}\left(B^{\star} \cW^{\top} (\cW\cW^{\top})^{-1} W_j -  \beta_j\right).$$
	note that $X H_j = Z \Sigma^{1/2} H_j \in \real^n$ is a random vector with i.i.d. entries of mean zero, variance $\|\Sigma^{1/2}H_j\|^2$, and finite $\varphi$-th moment by Assumption \eqref{assume_rm}.
	Hence by the law of large numbers, we have that with high probability
	${\left\|Z v_j \right\|^2 } = { n \|v_j\|^2} \cdot (1 + \OO(p^{-\e}))$, for a small enough constant $\e$,
	which implies that
	\begin{align}
		\abs{\sum_{j=1}^t \|X H_j\|^2 - \exarg{X}{\sum_{j=1}^t \norm{X H_j}^2}} \lesssim p^{-c} \cdot n \bignorm{\Sigma^{1/2} B^{\star} (U_{A}U_A^{\top} - \id_{t\times t})}^2. \label{eq_gA_err1}
	\end{align}

@@

		\abs{\normFro{\cE}^2 - \ex{\normFro{\cE}^2}} \le p^{-c_{\infty}} \ex{\normFro{\cE}^2}. \label{eq_gA_err3}

@@

		&\le p^{-c} \cdot n \bignormFro{\Sigma^{1/2} B^{\star}}^2 + p^{-1/2 + \e}\cdot \sigma^2 n \cdot t. \label{eq_g_gap}

### 2020-09-26 arxiv/setup.tex
wip

\subsection{Summary of Results}
\label{sec_prelim}

@@

\paragraph{Organizations.}
The rest of this paper is organized as follows.
In Section \ref{set_prelim}, we present a preliminary of hard parameter sharing and random matrices. We describe several well-known facts.
In Section \ref{sec_example}, we describe illustrative examples of our technical results and discuss their implications.
In Section \ref{sec_same}, we consider the same covariates setting and show the generalization error of hard parameter sharing.
In Section \ref{sec_diff}, we consider the different covariates setting and show the bias and variance asymptotics using random matrix theory.

@@

%Second, we assume that the spectral norm of $\Sigma_k$ is bounded by a fixed constant, for every $k = 1,\dots, t$.
%This is without loss of generality since we can always rescale $\Sigma_k$ so that its spectral norm satisfies the condition.
Finally, for the random noise component, we assume that all of its moments exist.
More precisely, there exists a fixed function $C(\cdot) : \mathbb{Z} \rightarrow \real^+$ such that for any $a = 1, 2, \dots, \infty$, we have that
\begin{align}\label{assmAhigh2}
	\ex{\abs{\varepsilon_{j}^{(i)}}^a} \le C(a), \text{ for any } 1\le i\le t \text{ and } 1\le j\le n_i.
\end{align}
Hence, for any value $\varphi > 4$, we get that Fact \ref{lem_minv} holds for $\varepsilon^{(i)}$, for all $i = 1, 2, \dots, t$.
Let $\e$ be an fixed value within $(0, 1/2)$.

>>>>

\subsection{Summary of Results}

@@

\paragraph{Organizations.}
The rest of this paper is organized as follows.
In Section \ref{sec_prelim}, we present a preliminary of hard parameter sharing and random matrices. We describe several well-known facts.
In Section \ref{sec_example}, we describe illustrative examples of our technical results and discuss their implications.
In Section \ref{sec_same}, we consider the same covariates setting and show the generalization error of hard parameter sharing.
In Section \ref{sec_diff}, we consider the different covariates setting and show the bias and variance asymptotics using random matrix theory.

@@

%Second, we assume that the spectral norm of $\Sigma_k$ is bounded by a fixed constant, for every $k = 1,\dots, t$.
%This is without loss of generality since we can always rescale $\Sigma_k$ so that its spectral norm satisfies the condition.
Finally, for the random noise component, we assume that all of its moments exist.
More precisely, there exists a fixed function $C(\cdot) : \mathbb{Z} \rightarrow \real^+$ such that for any $a = 1, 2, \dots, \infty$, we have that
\begin{align}\label{assmAhigh2}
	\ex{\abs{\varepsilon_{j}^{(i)}}^a} \le C(a), \text{ for any } 1\le i\le t \text{ and } 1\le j\le n_i.
\end{align}
Hence, for any value $\varphi > 4$, we get that Fact \ref{lem_minv} holds for $\varepsilon^{(i)}$, for all $i = 1, 2, \dots, t$.
Let $\e$ be a small enough fixed value and let $c_{\infty}$ be any fixed value within $(0, 1/2-\e)$.
We have that Fact \ref{lem_minv} holds for $\varepsilon^{(i)}$ where $c_{\varphi}$ becomes $c_{\infty}$ instead.
